---
title: Utilization and analysis of Fiji benthic data
author: Allen Coral Atlas Partnership
output:
  pdf_document:
    toc: yes
---

## Background
The [Allen Coral Atlas](http://allencoralatlas.org) partnership uses high-resolution satellite imagery and advanced analytics to map and monitor the world's coral reefs, creating unprecedented global coverage. As the Atlas develops maps of benthic habitat and reef geomorphology regionally and then globally, the field engagement component of the partnership seeks to identify and enable users of the Atlas to achieve conservation results (e.g., through marine spatial planning or other efforts).

These data can can hopefully be useful in planning and answering ecological research questions concerning marine and coastal habitats. These data, when combined with local knowledge, practioners's expertise, and other datasets can help guide and improvement conservation and management.

## Data requirements
In the accompanying GIS folder, you will find a geodatabase that contains all the required data if you would rather conduct the analysis in a GIS software you are more familiar with (see the separate tutorial for ArcGIS and QGIS).

After completing this and you want to perform similar analyses at a different geographic location, you can obtain data for other locations and datasets from the following locations:
* **Coral:** [Allen Coral Atlas](http://allencoralatlas.org)
* **Mangrove:** [Global Mangrove Watch](https://data.unep-wcmc.org/datasets/45)
* **Landmass:** [GADM](https://gadm.org/index.html)
* **Exclusive economic zone:** [Marine Regions](http://www.marineregions.org/)
* **Protected Areas:** [World Database on Protected Planet](https://www.protectedplanet.net/)


## Process
You will rely on many R packages to achieve the objectives. For those not installed, we will use the "pacman" package to install them now

```{r}
rm(list = ls()) # This will make sure you begin with a clean environment

if (!require("pacman")) install.packages("pacman") # installs the "pacman" function if not already installed

# load and install needed packages
pacman::p_load(dplyr,sf,sp,rgeos,rgdal,scales,raster,cartography,ggmap,ggplot2,wdpar,prioritizr,tmap,furrr,lubridate,maptools)

```

```{r}
# Load the Fiji area of interest boundary box
fiji_aoi <- st_read(dsn = "C:\\Users\\free\\Desktop\\Atlas_R\\GIS\\management", layer = "aoi") %>%
  st_transform("+proj=longlat +ellps=WGS84 +pm=jakarta +no_defs")

# Going to be using the same coordinate reference system in other layers
crs <- crs(fiji_aoi)
crs

# Now we will load adminstrative boundaries from global datasets and filter for specifically Fiji
fiji <- st_read(dsn = "C:\\Users\\free\\Desktop\\Atlas_R\\GIS\\administrative", layer = "countries") %>%
  dplyr::filter(NAME_0 =="Fiji") %>%
  st_transform(crs)
fiji_eez <- st_read(dsn = "C:\\Users\\free\\Desktop\\Atlas_R\\GIS\\administrative", layer = "eez") %>%
  dplyr::filter(ISO_SOV1=="FJI") %>%
  st_transform(crs)

# Marine and coastal benthic data are crucial for ecological and resource management questions, so we should add them
atlas_fiji <- st_read(dsn = "C:\\Users\\free\\Desktop\\Atlas_R\\GIS\\ecological", layer = "atlas_fiji") %>%
  st_transform(crs) %>%
  st_crop(fiji_aoi)

fiji_mangrove1996 <- st_read(dsn = "C:\\Users\\free\\Desktop\\Atlas_R\\GIS\\ecological", layer = "mangrove1996") %>%
  st_transform(crs) %>%
  st_crop(fiji_aoi)

fiji_mangrove2016 <- st_read(dsn = "C:\\Users\\free\\Desktop\\Atlas_R\\GIS\\ecological", layer = "mangrove2016") %>%
  st_transform(crs) %>%
  st_crop(fiji_aoi)

# Calculate total mangrove coverage in the AOI
fiji_mang_area1996 <- sum(st_area(fiji_mangrove1996))/1000000 # calculate km^2
fiji_mang_area2016 <- sum(st_area(fiji_mangrove2016))/1000000 # calculate km^2

# Calculate the difference between 1996 and 2016
fiji_mang_loss <- fiji_mang_area1996 - fiji_mang_area2016
```

## Quick check-in
Did Fiji have more managroves in 1996 or 2016?

```{r}
# Examining the major benthic categories from the Atlas data
# Begin by taking a look at the different benthic classes in the data
unique(atlas_fiji$benthic_cl)

# Let's delve into three specific classes: (1) Coral, (2) Coral/algae, and (3) Seagrass

fiji_coral <- atlas_fiji[atlas_fiji$benthic_cl == "Coral",]
View(fiji_coral)
fiji_coral_area <- sum(st_area(fiji_coral))/1000000 # calculate km^2
fiji_coral_area
summary(fiji_coral)

fiji_coral_algae <- atlas_fiji[atlas_fiji$benthic_cl == "Coral/Algae",]
View(fiji_coral_algae)
fiji_coral_algae_area <- sum(st_area(fiji_coral_algae))/1000000 # calculate km^2
fiji_coral_algae_area

fiji_seagrass <- atlas_fiji[atlas_fiji$benthic_cl == "Seagrass",]
View(fiji_seagrass)
fiji_seagrass_area <- sum(st_area(fiji_seagrass))/1000000 # calculate km^2
fiji_seagrass_area

# Calculate area for all benthic types
fiji_atlas_area <- atlas_fiji %>%                        # start with the Fiji benthic dataset
  dplyr::group_by(benthic_cl) %>%                        # aggregate by the benthic class
  dplyr::summarise(total_area = sum(area_sqkm, na.rm=T)) # create new field 'total' and fill it with the summed total area in km^2
```

## Quick check-in
Which habitat has the greatest extent coverage around Viti Levu? 

## Mapping
Knowning the total extent for benthic habitats are a great first step. For planning it is useful to know where the habitats are. This can help with understanding how they relate to each other and could be used for a marine spatial plan, for example.

```{r}
## Where are the habitats -- lets map them
### Atlas benthic classes across Fiji

ggplot() + 
  geom_sf(data = fiji, aes(fill = "grey85", col = "black", size = 0.3)) # national administrative boundary
  geom_sf(data = fiji_mangrove2016, aes(fill="#3a5d43")) + # plotting recent mangrove data
  geom_sf(data = atlas_fiji$benthic_cl, fill=atlas_fiji$benthic_cl) + # plotting Atlas benthic data
  coord_sf(xlim=c(-177,-179),ylim=c(-18,-16)) # center around Viti Levu
```

## Habitat change
Habitat loss is a major driver of biodiversity loss. The mangrove data can provide a temporal look at habitat change--how habitat is loss, but also if an area has been gained. These data will not detail when change occurred, the reason a loss (e.g., deforestation, invasive species), habitat quality (e.g., diseased), nor even provide how a gain was obtained (e.g., reforesation). As we have seen, above, they can say if overall there were gains or losses. Now let's examine the locations of those losses and gains.

We will do so using raster data (images).

```{r}
# set wd
wd1 <- setwd("C:\\Users\\free\\Desktop\\Atlas_R\\GIS\\ecological")

# load the mangrove rasters
mang_rast1996 <- raster("mang_aoi1996.tif")
mang_rast2016 <- raster("mang_aoi2016.tif")

freq(mang_rast1996) # how many times mangroves appeared in 1996
freq(mang_rast2016) # how many times mangroves appeared in 2016

# calculate the loss and gain values
mang_loss <- overlay(mang_rast1996, mang_rast2016, fun = function(mang_rast1996, mang_rast2016){mang_rast1996-mang_rast2016})
mang_gain <- overlay(mang_rast2016, mang_rast1996, fun = function(mang_rast2016, mang_rast1996){mang_rast2016-mang_rast1996})


plot(mang_loss) # see locations that once had mangrove
plot(mang_gain) # locations in 2016 that were not in 1996

# Mangrove extent 1996 vs. 2016
ggplot() +
  geom_raster(mang_rast1996, mapping = aes(col="216A12")) +
  geom_raster(mang_rast2016, mapping = aes(col="68B159"))

# Mangrove change
ggplot() +
  geom_raster(data=man_loss, mapping = aes(col = "CD796C")) + # areas of loss will be plotted in red
  geom_raster(data=mang_gain, mapping = aes(col = "509866")) # areas of gain will be plotted in green

```

### Fiji protected areas

```{r}
# WDPA marine classified
# load Fiji WDPA data --> clean and only marine areas
fiji_wdpa_raw <- wdpa_fetch("Fiji")
fiji_wdpa_clean <- wdpa_clean(fiji_wdpa_raw)

# Subset for the marine and coastal areas
fiji_marine_pa <- fiji_wdpa_clean %>%
  filter(GEOMETRY_TYPE=="POLYGON" & MARINE=="marine")
 
View(fiji_marine_pa)

# How much each protected area 
mpa_type_area <- fiji_marine_pa %>%
  dplyr::group_by(DESIG) %>%
  dplyr::summarise(total_area = sum(REP_M_AREA, na.rm=T))

# How much each year were protected areas added
mpa_year_area <- fiji_marine_pa %>%
  dplyr::group_by(STATUS_YR) %>%
  dplyr::summarise(total_area = sum(REP_M_AREA, na.rm=T))

plot(mpa_year_area$STATUS_YR,mpa_year_area$total_area,xlab="Year", ylab="Area (km^2)")

# Calculate total marine protected area
fiji_mpa_total <- sum(fiji_marine_pa$REP_M_AREA, na.rm = TRUE)

# Calculate the percentage of the no take areas across Fiji
fiji_marine_pa$no_tk_pct <- (fiji_marine_pa$NO_TK_AREA / fiji_marine_pa$REP_M_AREA) * 100

# Is the percentage of no take increasing over time?
mpa_notk_yr <- fiji_marine_pa %>%
  dplyr::group_by(STATUS_YR) %>%
  dplyr::summarise(mean_notk = mean((NO_TK_AREA / REP_M_AREA) * 100, na.rm=T))

plot(mpa_notk_yr$STATUS_YR,mpa_notk_yr$mean_notk,xlab="Year",ylab="")
```

##Future
Coastal and marine habitats are complex systems. They provide nursery habitats and feeding grounds for many species, protect coastal communities, as well as livelihoods for humans. By combining fisheries datasets, we can provide further insights on management decisions. We will use [RAM Legacy](https://www.ramlegacy.org/) and  [SeaAroundUs](http://www.seaaroundus.org/) to find key commercial catch species. [AquaMaps](https://www.aquamaps.org/) will provide data on species distributions to calculate and visualize species richness indexes. Further, we can leverage the open-source [Global Fishing Watch](https://globalfishingwatch.org/) data for its fishing vessel and effort data to visualize where and when vessels activity could impact habitat health and fisheries status.

Before moving towards the third section, we will synthesize and interpret the relationships between current protected areas and the marine and coastal habitats. Questions that will be addressed will be:
1. How much of each habitat is in each protected area?
2. What is the representativeness of each habitat in each protected area?
3. Are the protected benthic habitats more concentrated or distributed geographically?
4. Have protected benthic habitats been evenly protected over time?

Additionally, to identify positive solutions for conservation planning, thus benefiting humans and the environment. Marine spatial planning is a systamatic tool that incorporates a framework to ingest ecological, social, economicand polical data to find those positive solutions. We will use the "prioritizr" package to idenitfy a range of solutions that are representative and efficiently meet the desired objectives.
