---
title: "Oceana Illegal Fishing and Transparency"
author: "Brian Free"
date: "6/22/2022"
output:
  pdf_document:
    toc: yes
  html_document:
    df_print: paged
    toc: yes
  word_document:
    toc: yes
html_document:
  df_print: paged
  toc: yes
---

***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo}
# Clean environment by removing everything from the list
rm(list = ls())
```

```{r workspace}
#############################################
#############################################
##### 1. Get workspace ready

# Install and load required packages 
if (!require("pacman")) install.packages("pacman")
pacman::p_load(cowplot,
               devtools,
               dplyr,
               ggalt,
               ggplot2,
               gissr, # remotes::install_github("skgrange/gissr")
               grid,
               lubridate,
               plyr,
               raster,
               rgdal,
               rgeos,
               rnaturalearth, # use devtools::install_github("ropenscilabs/rnaturalearth") if packages does not install properly
               scales,
               sf,
               sp,
               stringr,
               tidyr)
```

```{r eval = FALSE}
# Check package versions
sessionInfo()

# Update packages if needed
# update.packages()
```

# Introduction
Oceana obtained fishing vessel data for the [Raymi](https://www.marinetraffic.com/en/ais/details/ships/shipid:168217/mmsi:224892000/imo:9024619/vessel:RAYMI). These data included its position during a 2-month period in 2017 and the activities it took during that voyage. From these data, this document will demonstrate how long the vessel was at voyage, where the vessel spent its voyage, and how many hours the voyage took during its activities.

# Methodology
These data were cleaned and analyzed in R to provide trends and insights into the Raymi's during June and July 2017.

Below outlines the following steps taken to analyze and summarize the data:

* Load data
  + Boundary data
  + Track data
  + Vessel data
* Inspect data
  + Track data
  + Vessel data
    1. Anchorage
    2. Fishing
* Clean data
  + Track data
  + Vessel data
* Transform data
  + Vessel data
    1. Time
    2. Event type
* Analyze data
  + Days a sea
  + Total event hours
  + Daily event hours
  + Monthly event hours
* Visualize data
* Patterns / insights
* Export data
  + Track
  + Vessel
  + Summary data
  + Spatial
  + Figures

***

# Set working directory and other key directories

```{r directory}
#############################################
#############################################
##### 2. Define presets and data dictionary

# Pre-set values (in this case the year and fishing classes are the most useful)

# Define data directory (since this is in R Project, pathname is simplified)
data <- "data"

# Define visuals directory (for exporting maps)
plotdir <- "figures"

# Define geopackage
geopackage <- "data/oceana_spatial.gpkg"

# map theme
theme1 <- theme(axis.title=element_blank(),
                # Gridlines
                panel.grid.major = element_blank(),
                panel.grid.minor = element_blank(),
                panel.background = element_blank(),
                axis.line = element_line(colour = "black"))
```


# Load data
The Raymi's vessel activity and track datasets were loaded from the data directory. Boundary data for the world and Peru were pulled from the [rnaturalearth R package](https://cran.r-project.org/web/packages/rnaturalearth/README.html). Exclusive Economic Zone boundary data came from [Marine Regions](https://marineregions.org/downloads.php). It should be noted these are older data, as the most current version is 11 from 2019.

```{r load, echo = T, results = 'hide'}
#############################################
#############################################
##### 3. Read in vessel and boundary data

# Read data
## Vessel data
vessels <- read.csv(paste(data, "events.csv", sep = "/"))

## Track data
tracks <- read.csv(paste(data, "tracks.csv", sep = "/"))

# World and country data
world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")
peru <- rnaturalearth::ne_countries(country = "Peru", scale = "medium", returnclass = "sf")

# EEZ data
eez <- st_read(dsn = data, layer = "World_EEZ_v8_2014")
```

# Inspect data
It is common for disruptions and other issues to the Global Positioning System to affect the record of the logging coordinates. Thus, plots of the latitudes and longitudes were run to determine if there were any outliers. For instance, there was a record of what appeared to be a latitude of 100, which is not possible since latitudes run from -90 to 90. Aside from that, there was a single observation that appeared to not follow the majority of records. Longtitude also had two values that appeared to be outliers.

Vessel events data had two different types recorded: fishing and anchoring.
```{r inspect}
#############################################
#############################################
##### 4. Inspect the data
names(tracks)
plot(tracks$lat) # can see that there is a latitude with a value near 100, which is not possible, also an outlier likely with a single observation well below -50
plot(tracks$lon) # similarly two values appear to be outliers for the longitude

# vessel events data
head(vessels)
unique(vessels$event_type) # only two events are fishing and anchoring
```

# Clean data
To have the best available data for analysis, the outliers were removed from the tracks dataset.

```{r clean}
#############################################
#############################################
##### 5. Clean data for preparation

# remove any NA values and what appear to be outlier observations
tracks_clean <- tracks %>%
  # remove any longitude and latitude values that are NA
  filter(!is.na(lat) & # latitude
         !is.na(lon) & # longitude
         !(lat >= 90) & # vessels should not be above 90 degrees
         !(lat <= -50)) %>% # vessels should not be below -50 degrees
  tidyr::separate(timestamp, into=c("date", "time"), sep=" ", remove=F, convert = T) %>%
  dplyr::arrange(X)

View(tracks_clean)

plot(tracks_clean$lat) # these plots make more sense
plot(tracks_clean$lon) # these plots make more sense

rm(tracks) # no longer will be needed
```

# Transform data
For the vessel data, new fields were created for latitude, longitude, and time respectively. Additionally, from the event start and end timestamps, a new field contained total duration in hours for each event.
```{r vessel transform}
#############################################
#############################################
##### 6. Transform vessel data
vessels_clean <- vessels %>%
  # Separate lat/long/start
  tidyr::separate(latlonstart, into=c("lat", "long", "date", "time"), sep=" ", remove=F, convert = T) %>% 
  # Build event start time
  mutate(event_start=paste(date, time)) %>% 
  # Calculate event duration
  mutate(event_duration_hr = difftime(event_end, event_start, units = "hours")) %>% 
  # Simplify
  dplyr::select(vessel_1_id, event_type, event_start, event_end, event_duration_hr, lat, long) %>%
  # Recode activity
  dplyr::mutate(event_type = recode(event_type,
                                    "anchorage" = "Anchorage",
                                    "fishing" = "Fishing"))

rm(vessels)
```

```{r event transform}
fishing <- vessels_clean %>%
  dplyr::filter(event_type == "Fishing") %>%
  st_as_sf(., coords = c("long", "lat"), crs = 4326)

fishing

anchorage <- vessels_clean %>%
  dplyr::filter(event_type == "Anchorage") %>%
  st_as_sf(., coords = c("long", "lat"), crs = 4326)

anchorage
```

# Visualize data
```{r}
#############################################
#############################################
##### 7. Visualize data
vessel_world_pos <- ggplot() +
  # Plot EEZ
  geom_sf(data = eez, linetype = "dashed", fill = NA, color = "grey10", size = 0.3) +
  # Plot countries
  geom_sf(data = world, fill = "grey80", color = "white", size = 0.4) +
  geom_sf(data = peru, fill = "grey60", color = "white", size = 0.4) +
  # Plot tracks
  geom_path(data = tracks_clean, mapping=aes(x = lon, y = lat), linetype = "solid", color = "#696969") +
  # Plot points
  geom_point(data = vessels_clean, mapping=aes(x=long, y=lat, color=event_type), size = 0.5) +
  # Crop
  coord_sf(xlim = c(-65, -85), ylim = c(-25, -10)) +
  # Labels
  # Obtained ship name from searching for the MMSI value (https://www.marinetraffic.com/en/ais/details/ships/shipid:168217/mmsi:224892000/imo:9024619/vessel:RAYMI)
  labs(title = paste0("Raymi\n",vessels_clean$vessel_1_id[1]), 
       caption = "Vessel tracks for Raymi (MMSI 224892000) during a
       fishing trip lasting from 2017-06-01 to 2017-08-01.\nData source: Global Fishing Watch") +
  # Legend
  scale_color_manual(name="Activity", values=c("red", "green")) +
  # Theme
  theme_bw() +  theme1

vessel_world_pos
```

# Patterns / trends
The Raymi for June and July 2017 spent its voyage to the west of Peru and Chile's exclusive economic zone. During its 60-day voyage it landed at a port in Peru for 131 hours before setting a sea for more fishing. In total it spent more than 921 hours fishing. Since activity hours are summarized between time stamps and not at the end of the day, total time spent on an activity during each day a sea can have some error. Given that, the longest times fishing were for 62 hours starting on July 8th and 55 hours from July 30th. The four least active days were in June, whereas 9 of top 10 most active record fishing activities were in July. July had more than double logged fishing hours than June, nearly 625 hours compared to 297 hours. Again, the Raymi was at port for 131 hours, which occurred in June, but that still is 150 hours shorter than the time in July. Should be noted that July is a day longer than June -- though that extra 24 hours cannot explain the 150 hour difference. Fishing time appears to oscillate. This could be due to changes in climatic or fish dynamics, but perhaps also happens so crews can recover. With very long fishing followed by shorter fishing time. The crew logged approximately 15,250 kilometers during the voyage.


```{r open ended section}
#############################################
#############################################
##### 8. Summary statistics

# Days at sea
days_asea <- tracks_clean %>%
  group_by(date) %>%
  summarise(Event = mdy(date),
            Days = Event - lag(Event)) %>%
  dplyr::filter(Days > 0)

asea_time <- nrow(days_asea)
print(paste(asea_time, "days at sea"))

# event hours
event_hours <- vessels_clean %>%
  dplyr::group_by(event_type) %>%
  dplyr::summarise(hours = sum(event_duration_hr))
head(event_hours)

# Event type hours by month
fishing_date <- vessels_clean %>%
  tidyr::separate(event_end, into = c("date", "time"), sep = " ", remove=F, convert = T) %>%
  # Build event start time
  mutate(event_start=paste(date, time)) %>%
  # Simplify
  dplyr::select(vessel_1_id, event_type, date, event_duration_hr, lat, long) %>%
  # Group by date and event type
  dplyr::group_by(date,
                  event_type) %>%
  # Calculate total hours by event per day
  dplyr::summarise(hours = sum(event_duration_hr)) %>%
  tidyr::separate(hours, into = c("time", "unit"), sep = " ", remove = F, convert = T) %>%
  # Simplify
  dplyr::select(-unit) %>%
  dplyr::arrange(desc(time))
View(fishing_date)

raymi_fishing_date <- ggplot() +
  geom_point(data = fishing_date, aes(x = date, y = time, color = event_type)) +
  # Legend
  scale_color_manual(name="Activity", values=c(c("#87468B", "#BFCE67"))) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text.x = element_text(angle = 90)) +
  labs(title = paste0("Raymi\n",vessels_clean$vessel_1_id[1]), 
       caption = "Total hours for each event per day for Raymi (MMSI 224892000)
       during a fishing trip lasting from 2017-06-01 to 2017-08-01.\nData source: Global Fishing Watch") +
  xlab("Date") +
  ylab("Total hours")
raymi_fishing_date

month_hours <- fishing_date %>%
  # Convert date to date class
  dplyr::mutate(date = as.Date(date)) %>%
  # Calculate month
  dplyr::mutate(month = lubridate::month(date)) %>%
  # Group by month and event
  dplyr::group_by(month,
                  event_type) %>%
  dplyr::summarise(month_hours = sum(hours))
head(month_hours)

# distance traveled
track_points_df <- tracks_clean %>% 
  mutate(group="track") %>% 
  dplyr::select(group, lon, lat)
track_points_sf <- track_points_df %>% 
  sf::st_as_sf(coords=c("lon", "lat"), crs=sf::st_crs(world))
track_line_sf <- track_points_sf %>% 
  dplyr::group_by(group) %>%
  dplyr::summarise(do_union = FALSE) %>%
  sf::st_cast("MULTILINESTRING") 
distance <- sf::st_length(track_line_sf)

print(paste(round(distance, 0), "m travelled by the Raymi during June 2017 through July 2017"))
```

# Export data
## Tracks
```{r export}
#############################################
#############################################
##### 9. Export data
saveRDS(tracks_clean, file=file.path(data, "raymi_tracks_062017-072017.Rds"))
```

## Vessels
```{r}
saveRDS(vessels_clean, file=file.path(data, "raymi_events_062017-072017.Rds"))
```

## Summary data
```{r}
saveRDS(event_hours, file=file.path(data, "raymi_event_hrs_062017-072017.Rds"))
saveRDS(fishing_date, file=file.path(data, "raymi_fishing_date_062017-072017.Rds"))
saveRDS(month_hours, file=file.path(data, "raymi_mnth_hrs_062017-072017.Rds"))
```

## Geopackage
```{r}
st_write(obj = track_points_sf, dsn = geopackage, layer = "raymi_points", append = F)
st_write(obj = track_line_sf, dsn = geopackage, layer = "raymi_track", append = F)
st_write(obj = world, dsn = geopackage, layer = "world_countries", append = F)
st_write(obj = peru, dsn = geopackage, layer = "peru", append = F)
st_write(obj = eez, dsn = geopackage, layer = "eez", append = F)
```

## Map
### Save plots
```{r}
ggsave(vessel_world_pos, filename = file.path(plotdir, "peru_fishing_map.png"),
       width = 12, height = 8, units = "in", dpi = 600)
ggsave(raymi_fishing_date, filename = file.path(plotdir, "raymi_daily_activity_time.png"),
       width = 12, height = 8, units = "in", dpi = 600)
```

